Acceleration of plague outbreaks in the second pandemic\\\emph{Electronic \supp}

### 2020-06-18 supp.Rnw
more detail on link functions; three vs two bad fits

Robustly fitting phenomenological epidemic curves to small, noisy data
sets proved to be surprisingly challenging\footnote{See ``monsters in
  the basement'',
  \url{https://redpenblackpen.tumblr.com/post/145304820562/monsters-in-the-basement}.}. In
the course of developing the full model we experimented with a range
of optimizers --- specifically the BFGS and Nelder-Mead options for
R's \code{optim} function as well as R's \code{nlminb} function; we
also tried an approach that iterated back and forth between
Nelder-Mead and BFGS until convergence was achieved or the fit was
sufficiently stable. We chose \code{nlminb} because it gave the best
results (highest log likelihood) for point estimates. In general, we used the standard link functions proposed by Ma \emph{et al.} \cite{Ma+14}, but we found that the scaled-tangent link proposed by \cite{Ma+14} ($\eta = \tan((x/2-1) \cdot \pi/2)$) declined too slowly in the tails, leading to numerical instability; we replaced it with a more standard logit link ($\eta = \log(x/(1-x))$).

With this approach we failed to achieve good (convergent) fits for only two combinations of source and outbreak year, both for minor outbreaks (London bills for 1578, Canterbury wills for 1581); these cases are excluded from the tables and figures shown below.

@@

%%\clearpage
%%\bibliographystyle{vancouver}
%%\bibliographystyle{Science}
\bibliographystyle{pnas-new}
\bibliography{plague}
\addcontentsline{toc}{section}{References}

>>>>

Robustly fitting phenomenological epidemic curves to small, noisy data
sets proved to be surprisingly challenging\footnote{See ``monsters in
  the basement'',
  \url{https://redpenblackpen.tumblr.com/post/145304820562/monsters-in-the-basement}.}. In
the course of developing the full model we experimented with a range
of optimizers --- specifically the BFGS and Nelder-Mead options for
R's \code{optim} function as well as R's \code{nlminb} function; we
also tried an approach that iterated back and forth between
Nelder-Mead and BFGS until convergence was achieved or the fit was
sufficiently stable. We chose \code{nlminb} because it gave the best
results (highest log likelihood) for point estimates.
In general, we used the standard link functions proposed by Ma \emph{et al.} \cite{Ma+14}, i.e. fitting most of the parameters (growth rate $r$, background mortality/wills rate $b$, final size $K$, Richards shape parameter $s$) on the log scale with a scaled-tangent link ($\eta = \tan(\pi/2 \cdot (x/2-1))$) for the initial condition $x_0$. However, we found that the scaled-tangent link declined too slowly in the tails, leading to numerical instability; we replaced it with a more standard logit link ($\eta = \log(x/(1-x))$).

With this approach we failed to achieve good (convergent) fits for only three combinations of source and outbreak year, all for minor outbreaks (London bills for 1578 and 1582, Canterbury wills for 1581); these cases are excluded from the tables and figures shown below.

@@

%%\clearpage
\bibliographystyle{pnas-new}
\bibliography{plague}
\addcontentsline{toc}{section}{References}
